Software Citation Principles

### 2016-07-21 sc-principles/software-citation-principles.tex
adding more about software/data differences

The process of creating the software citation principles began by adapting
the FORCE11 Data Citation Principles~\cite{data-citation-principles}. These
were then modified based on discussions of the FORCE11 Software Citation
Working Group (see Appendix~\ref{app:wg_members} for members), information from
the use cases in \S\ref{sec:use_cases}, and the related work in
\S\ref{sec:related_work}.
The adaptations were made because software, while similar to data in terms of
not traditionally having been cited in publications, is also different than
data. Software can be used to express or explain concepts, it is typically
updated more frequently than data\footnote{Datasets can of course be frequently
updated, but each update likely represents the findings of a new or repeated
study.}, and it is executable.
\niemnote{Dan: can you add more here?}
Also, while software can be considered a type of data, the converse is not
generally true.

>>>>

The process of creating the software citation principles began by adapting
the FORCE11 Data Citation Principles~\cite{data-citation-principles}. These
were then modified based on discussions of the FORCE11 Software Citation
Working Group (see Appendix~\ref{app:wg_members} for members), information from
the use cases in \S\ref{sec:use_cases}, and the related work in
\S\ref{sec:related_work}.
The adaptations were made because software, while similar to data in terms of
not traditionally having been cited in publications, is also different than
data. While software can be considered a type of data, the converse is not
generally true.  In addition:

\begin{itemize}
\item Software can be used to express or explain concepts, unlike data.
\item Software is updated more frequently that papers or data\footnote{Datasets can of course be frequently updated, but each update likely represents the findings of a new or repeated study.}.
\item Software is executable, unlike papers or data.
\item Software suffers from a different type of bit rot than data --- software must be constantly maintained so that it continues to function as both the hardware and software environments that it is used on changes, as bugs are found and fixed, and as user requirements demand new features and capabilities.
\item Software is frequently built to use other software, leading to complex dependencies, and these dependent software packages also are frequently changing.
\item Software is generally smaller than data, so a number of the storage and preservation constraints on data don't apply to software.
\item Software teams can be very large, and very multidisciplinary, and there can be many varied roles that are overlapping and sometimes ambiguous with respect to contributions, and which contributions rise to a level equivalent to paper authorship.
\item The lifetime of software is generally not as long as that of data.
\end{itemize}
